---
title: "Visualizations V1"
subtitle: "LLM Bias"
author: "Lasse Hyldig Hansen"
date: "2024-19-01"
output: 
  html_document:
    toc: true
---

## Loading data

```{r}
df_gender_tot = read_csv("count_vis_data/df_gender_tot.csv")
df_racial_tot = read_csv("count_vis_data/df_racial_tot.csv")
```

## Packages

```{r}
library(tidyverse)
```


## Preprocssing 

### only for window == 50

```{r}
df_racial_tot2 = df_racial_tot %>% rename("keyword" = "racial_key") %>% 
                 mutate(
                   percentage = ifelse(is.nan(percentage), 0, percentage)
                 )
df_gender_tot2 = df_gender_tot %>% rename("keyword" = "gender_key") %>% 
                 mutate(
                   percentage = ifelse(is.nan(percentage), 0, percentage)
                 )


df_racial_tot3_perc = df_racial_tot2 %>%
    filter(window == 50) %>% 
    group_by(medical_key, keyword) %>%
    summarize(
      count = sum(count)
    ) %>%
    ungroup() %>% 
    group_by(medical_key) %>%
  mutate(
      total_count = sum(count),
      data_perc = (count/total_count) * 100
    )

df_gender_tot3_perc = df_gender_tot2 %>%
    filter(window == 50) %>% 
    group_by(medical_key, keyword) %>%
    summarize(
      count = sum(count)
    ) %>%
    ungroup() %>% 
    group_by(medical_key) %>%
  mutate(
      total_count = sum(count),
      data_perc = (count/total_count) * 100
    )

df_gender_racial_tot_perc = rbind(df_racial_tot3_perc, df_gender_tot3_perc)

df_racial_tot3_perc$keyword = factor(df_racial_tot3_perc$keyword, 
                                                levels = c(
                                                            "tot_disease_counts",
                                                            "white/caucasian", 
                                                                        "black", 
                                                                        "asian",
                                                                        "hispanic/latino",
                                                                        "native american/alaskan native",
                                                                        "pacific islander"
                                                           ))

df_gender_tot3_perc$keyword = factor(df_gender_tot3_perc$keyword, 
                                                levels = c("tot_disease_counts",
                                                                        "female",
                                                                        "male"))

df_gender_racial_tot_perc$keyword = factor(df_gender_racial_tot_perc$keyword, 
                                                levels = c(
                                                           "tot_disease_counts", "white/caucasian", 
                                                                        "black", 
                                                                        "asian",
                                                                        "hispanic/latino",
                                                                        "native american/alaskan native",
                                                                        "pacific islander",
                                                                        "female",
                                                                        "male"))

```

```{r}
# Now create the plot

### adding total percentages
sum_count = sum(df_gender_tot3_perc$count)

df_gender_tot3_perc$perc_of_total = (df_gender_tot3_perc$total_count/sum_count)*100

# Create a custom label for faceting
df_gender_tot3_perc <- df_gender_tot3_perc %>%
  mutate(facet_label = paste0(round(total_count / 1000, 1), 
                              " K (", 
                              round(perc_of_total, 2), "%)"))



### adding total percentages
sum_count = sum(df_racial_tot3_perc$count)

df_racial_tot3_perc$perc_of_total = (df_racial_tot3_perc$total_count/sum_count)*100

# Create a custom label for faceting
df_racial_tot3_perc <- df_racial_tot3_perc %>%
  mutate(facet_label = paste0(round(total_count / 1000, 1), 
                              " K (", 
                              round(perc_of_total, 2), "%)"))
```

### Bar visualizations 

```{r}
# Now create the plot
p1_stacked_bar = ggplot(data = df_gender_tot3_perc, aes(x = medical_key, y = data_perc, fill = keyword)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  facet_grid(reorder(facet_label, -total_count) ~ ., scales = "free_y",  switch = "y") +
  labs(title = "",
       x = "",
       y = "",
       fill = "") +
  scale_fill_manual(values = rev(c("#344e41", "#9b2226", "#DCDBA8")),
                    labels = c("No Disease Mention", "Female", "Male")) +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "right",
        strip.text.y.left = element_text(angle = 0),
        panel.spacing.x = unit(0, "lines"),
        panel.spacing.y = unit(0.1, "cm"),
        legend.margin = margin(t = -0.2, unit = "lines"),
        plot.margin = unit(c(-0.5, 0.2, -0.5, 0.2), 'lines'),
        legend.text = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 12, color = "black", face = "bold"),
        strip.text = element_text(size = 12, color = "black"),
        axis.text.x = element_text(size = 10, color = "black"),
        #strip.text = element_text(angle = 0, size = 10, face = "bold", margin = margin(t = 5, b = 5, l = 5, r = 5)),
        strip.background = element_rect(fill = "white", color = "black", size = 1)) +
  scale_y_continuous(expand = c(0,0))


p2_stacked_bar = ggplot(data = df_racial_tot3_perc, aes(x = medical_key, y = data_perc, fill = keyword)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  facet_grid(reorder(facet_label, -total_count) ~ ., scales = "free_y",  switch = "y") +
  #facet_grid(facet_label ~ ., scales = "free_y",  switch = "y") +
  labs(title = "",
       x = "",
       y = "",
       fill = "") +
  scale_fill_manual(values=rev(c("#bc6c25", "#344e41", "#540b0e", "#9b2226", "#03045e", "#0077b6", "#DCDBA8")),
                    labels = c("No Disease Mention", "White", "Black", "Asian", "Hispanic", "Native", "Pac. Islander")) +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "right",
        strip.text.y.left = element_text(angle = 0),
        panel.spacing.x = unit(0, "lines"),
        panel.spacing.y = unit(0.1, "cm"),
        legend.margin = margin(t = -0.2, unit = "lines"),
        plot.margin = unit(c(-0.5, 0.2, -0.5, 0.2), 'lines'),
        legend.text = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 12, color = "black", face = "bold"),
        strip.text = element_text(size = 12, color = "black"),
        axis.text.x = element_text(size = 10, color = "black"),
        #strip.text = element_text(angle = 0, size = 10, face = "bold", margin = margin(t = 5, b = 5, l = 5, r = 5)),
        strip.background = element_rect(fill = "white", color = "black", size = 1)) +
  scale_y_continuous(expand = c(0,0))
```

## Making source visualizations
### only for stratifying by data-source

```{r}

df_racial_tot2 = df_racial_tot %>% rename("keyword" = "racial_key") %>% 
                 mutate(
                   percentage = ifelse(is.nan(percentage), 0, percentage)
                 )
df_gender_tot2 = df_gender_tot %>% rename("keyword" = "gender_key") %>% 
                 mutate(
                   percentage = ifelse(is.nan(percentage), 0, percentage)
                 )

df_racial_tot3_perc = df_racial_tot2 %>%
    filter(window == 50) %>%
    group_by(medical_key, keyword, source) %>%
    summarize(
      count = sum(count)
    ) %>%
    ungroup() %>% 
    group_by(medical_key, source) %>%
  mutate(
      total_count = sum(count),
      data_perc = (count/total_count) * 100
    )

df_gender_tot3_perc = df_gender_tot2 %>%
    filter(window == 50) %>% 
    group_by(medical_key, keyword, source) %>%
    summarize(
      count = sum(count)
    ) %>%
    ungroup() %>% 
    group_by(medical_key, source) %>%
  mutate(
      total_count = sum(count),
      data_perc = (count/total_count) * 100
    )

df_gender_racial_tot_perc = rbind(df_racial_tot3_perc, df_gender_tot3_perc)

df_racial_tot3_perc$keyword = factor(df_racial_tot3_perc$keyword, 
                                                levels = c(
                                                            "tot_disease_counts",
                                                            "white/caucasian", 
                                                                        "black", 
                                                                        "asian",
                                                                        "hispanic/latino",
                                                                        "native american/alaskan native",
                                                                        "pacific islander"
                                                           ))

df_gender_tot3_perc$keyword = factor(df_gender_tot3_perc$keyword, 
                                                levels = c("tot_disease_counts",
                                                                        "female",
                                                                        "male"))

df_gender_racial_tot_perc$keyword = factor(df_gender_racial_tot_perc$keyword, 
                                                levels = c(
                                                           "tot_disease_counts", "white/caucasian", 
                                                                        "black", 
                                                                        "asian",
                                                                        "hispanic/latino",
                                                                        "native american/alaskan native",
                                                                        "pacific islander",
                                                                        "female",
                                                                        "male"))

```

### Adding total percentages for sources

#### For gender

```{r}
### adding total percentages
sum_count_sources = sum(df_gender_tot3_perc$count)

df_gender_tot3_perc$perc_of_total = (df_gender_tot3_perc$total_count/sum_count_sources)*100

## getting source counts 

df_gender_tot3_perc = df_gender_tot3_perc %>% 
  group_by(source) %>% 
  mutate(source_count = sum(count),
         source_perc = (source_count/sum_count_sources)*100)

# Create a custom label for faceting
df_gender_tot3_perc <- df_gender_tot3_perc %>%
  mutate(facet_label = paste0(source,
                              "\n",
                              round(source_count / 1000, 1), 
                              " K (", 
                              round(source_perc, 2), "%)"))
```

#### For race

```{r}
### adding total percentages
sum_count_sources = sum(df_racial_tot3_perc$count)

df_racial_tot3_perc$perc_of_total = (df_racial_tot3_perc$total_count/sum_count_sources)*100

## getting source counts 

df_racial_tot3_perc = df_racial_tot3_perc %>% 
  group_by(source) %>% 
  mutate(source_count = sum(count),
         source_perc = (source_count/sum_count_sources)*100)

# Create a custom label for faceting
df_racial_tot3_perc <- df_racial_tot3_perc %>%
  mutate(facet_label = paste0(source,
                              "\n",
                              round(source_count / 1000, 1), 
                              " K (", 
                              round(source_perc, 2), "%)"))
```

## Source visualizations

```{r}
# Create the updated stacked bar chart with facet labels
p1_stacked_bar_source = ggplot(data = df_gender_tot3_perc, aes(x = medical_key, y = data_perc, fill = keyword)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  labs(title = "",
       x = "",
       y = "",
       fill = "") +
scale_fill_manual(values = rev(c("#344e41", "#9b2226", "#DCDBA8")),
                    labels = c("No Disease Mention", "Female", "Male")) +  facet_grid(. ~ reorder(facet_label, -total_count)) +
    
  #facet_grid(. ~ facet_label, scales = "free") +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "bottom",
        #strip.text.y.left = element_text(angle = 0),
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(0.4, "cm"),
        legend.margin = margin(t = -1, unit = "lines"),  
        plot.margin = unit(c(-0.5, 0.5, 0.2, 0.2), 'lines'),
        legend.text = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 12, color = "black"),
        strip.text = element_text(size = 12, color = "black"),
        axis.text.x = element_text(size = 10, color = "black"),
        
        strip.text.y.left = element_text(angle = 0, size = 10, face = "bold", margin = margin(t = 5, b = 5, l = 5, r = 5)),
        strip.background = element_rect(fill = "white", color = "black", size = 1.5)) +
  scale_y_continuous(expand = c(0,0))

p2_stacked_bar_source = ggplot(data = df_racial_tot3_perc, aes(x = medical_key, y = data_perc, fill = keyword)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  labs(title = "",
       x = "",
       y = "",
       fill = "") +
  scale_fill_manual(values=rev(c("#bc6c25", "#344e41", "#540b0e", "#9b2226", "#03045e", "#0077b6", "#DCDBA8")),
                    labels = c("No Disease Mention", "White", "Black", "Asian", "Hispanic", "Native", "Pac. Islander")) +
  facet_grid(. ~ reorder(facet_label, -total_count)) +
    
  #facet_grid(. ~ facet_label, scales = "free") +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "bottom",
        #strip.text.y.left = element_text(angle = 0),
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(0.4, "cm"),
        legend.margin = margin(t = -1, unit = "lines"),  
        plot.margin = unit(c(-0.5, 0.5, 0.2, 0.2), 'lines'),
        legend.text = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 12, color = "black"),
        strip.text = element_text(size = 12, color = "black"),
        axis.text.x = element_text(size = 10, color = "black"),
        
        strip.text.y.left = element_text(angle = 0, size = 10, face = "bold", margin = margin(t = 5, b = 5, l = 5, r = 5)),
        strip.background = element_rect(fill = "white", color = "black", size = 1.5)) +
  scale_y_continuous(expand = c(0,0))
```


### I HAVE NOT DONE THE LOWER PART YET 

### I HAVE DATA FROM CODING INEQUITY, IN THE DATA FOLDER HERE 

## Adding data from prevelance 

```{r}
final_true_dist_2_ <- read_csv("data_coding_inequity/final_true_dist (2).csv")
final_true_dist_2_$...1 = NULL
```

### Cleaning it to fit my data

```{r}
unique(df_racial_tot3_perc$medical_key)
## removing diseases we don't have


final_true_dist_2_ = final_true_dist_2_ %>% filter(Condition != "Urinary tract infection" &
                                                   Condition != "Preeclampsia" &
                                                  Condition !=  "Chlamydia" ) 

final_true_dist_2_$medical_key = 
  c("rheumatoid arthritis", 
  "hepatitis b",
  "endocarditis", 
  "covid-19",
  # Dont have Urinary tract infection
  "bacterial pneumonia",
  "takotsubo cardiomyopathy",
  "essential hypertension",
  "type 2 diabetes", #leo used both t1 and t2
  ## Dont have preeclampsia
  "lupus", 
  "sarcoidoses",
  "colon cancer",  
  "syphilis",  
  ## Don't have chlamydia
  "hiv/aids", 
  "tuberculosis", 
   "multiple sclerosis", 
  "prostate cancer",   
   "multiple myeloma"
  )


```

```{r}

df_racial_tot2 = df_racial_tot %>% rename("keyword" = "racial_key") %>% 
                 mutate(
                   percentage = ifelse(is.nan(percentage), 0, percentage)
                 )
df_gender_tot2 = df_gender_tot %>% rename("keyword" = "gender_key") %>% 
                 mutate(
                   percentage = ifelse(is.nan(percentage), 0, percentage)
                 )

df_racial_tot3_perc = df_racial_tot2 %>%
    filter(window == 50 & keyword != "tot_disease_counts") %>% 
    mutate(keyword = ifelse(keyword %in% c("pacific islander", "native american/alaskan native"), "other", keyword)) %>% 
    group_by(medical_key, keyword) %>%
    summarize(
      count = sum(count)
    ) %>%
    ungroup() %>% 
    group_by(medical_key) %>%
  mutate(
      total_count = sum(count),
      data_perc = (count/total_count) * 100
    )

df_gender_tot3_perc = df_gender_tot2 %>%
    filter(window == 50 & keyword != "tot_disease_counts") %>% 
    mutate(keyword = ifelse(keyword %in% c("pacific islander", "native american/alaskan native"), "other", keyword)) %>% 
    group_by(medical_key, keyword) %>%
    summarize(
      count = sum(count)
    ) %>%
    ungroup() %>% 
    group_by(medical_key) %>%
  mutate(
      total_count = sum(count),
      data_perc = (count/total_count) * 100
    )

df_gender_racial_tot_perc = rbind(df_racial_tot3_perc, df_gender_tot3_perc)

df_gender_racial_tot_perc_wide = df_gender_racial_tot_perc %>%
  select(medical_key, keyword, data_perc) %>%
  pivot_wider(names_from = keyword, values_from = data_perc)
```

```{r}
colnames(df_gender_racial_tot_perc_wide) <- c("medical_key", "Asian", "Black", "Hispanic/Latino", "Other", "White", "Female", "Male")

final_true_dist_2_$Condition = NULL

colnames(final_true_dist_2_) <- c("Male", "Female", "Black","White",  "Hispanic/Latino", "Asian", "Other", "medical_key")
```


```{r}
df_gender_racial_tot_perc_wide = df_gender_racial_tot_perc_wide %>% filter(medical_key != "chagas disease" &
                                                   medical_key != "crohn’s disease" &
                                                  medical_key !=  "gastric cancer" &
                                                  medical_key !=  "ovarian cancer" &
                                                  medical_key !=  "ulcerative colitis") 
```


```{r}

df_gender_racial_tot_perc_wide$estimate = "Training Data"
final_true_dist_2_$estimate = "Real World Estimate"

together_plot = rbind(df_gender_racial_tot_perc_wide, final_true_dist_2_)

together_plot_long = together_plot %>% pivot_longer(2:8)

together_plot_long = together_plot_long %>% mutate(name = ifelse(name == "Hispanic/Latino", "Hispanic", name))


together_plot_long$name = factor(together_plot_long$name, 
                                                levels = c( "White", 
                                                                        "Black", 
                                                                        "Asian",
                                                                        "Hispanic",
                                                                        "Other",
                                                                        "Female",
                                                                        "Male"))

p4 = ggplot(together_plot_long, aes(x = value, y = estimate, fill = estimate)) +
  geom_bar(stat = 'identity', width = 0.9, color = "black") +  # increased bar width
  #coord_cartesian(clip = "off") + 
  scale_fill_manual(values = c("#bc6c25","#344e41")) +
#,
#                    labels = c("Training Data", "GTP-4 Estimated", "True")) + 
  labs(title = "",
       x = "Percentage",
       y = "",
       fill = "") +
  theme_minimal() +
  theme(
        strip.text.y.left  = element_text(angle = 0, size = 11),
        strip.text = element_text(angle = 0, size = 11, face = "bold"),
        strip.background = element_rect(fill = "white", color = "black", size = 0.5),
        panel.border = element_rect(fill = NA, color = "black", size = 0.5),
        legend.position = "none",
        axis.title.x = element_text(colour = "black", size = 10),
        axis.text.x = element_text(size=9, colour="black"),
        axis.text.y = element_text(size=10, , colour="black", hjust = 0)) +
  facet_grid(medical_key ~ name, scales = "free", switch = "y") +
  theme(
      panel.spacing.x = unit(0.5, "lines"),  # Reduced panel spacing
      panel.spacing.y = unit(0, "lines"),  # Reduced panel spacing
      plot.margin = unit(c(0, 0.5, 0.1 , 0), 'lines')) + 
  scale_x_continuous(limits = c(0, 100), expand = c(0,0),
                     breaks = c(0, 50, 90),
                     labels = c(0, 50, 100))



```

